---
author: "Svadrut Kukunooru"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(tidyverse)
library(lubridate)
```

## Assignment 2

- Replace the text YOUR NAME HERE in the YAML section with your name.
- Execute the code in the chunk below to read in the Lake Monona data. *(Note: this data set is different than that used in lecture.)*
    - If there is an error, check to see if there is a `data` directory two directories up that contains the data file.
- Post questions on our Piazza site, visit the Learning Center, or attend office hours if you have questions.
    
```{r read-data, echo = FALSE}
## The echo = FALSE argument to this chunk
##   means that the code will not appear in the output
##   but the code will run so that the data is in the session

## Read Lake Monona data
## Remove rows with missing data
monona = read_csv("../../data/lake-monona-winters-2021.csv") %>%
  drop_na()
```

### Problems

### 1

The following code makes a histogram of the `duration` variable in the Lake Monona data set.

```{r problem1, fig.height = 3}
ggplot(monona, aes(x=duration)) +
  geom_histogram(boundary=0, binwidth=10, color="black", fill="white")
```

Approximately how many winters had between 40 to 70 days where Lake Monona was at least 50% covered with ice?

### Response

11

### 2

Modify the code below so that in the resulting plot:

- one of the bins is centered at 70 days
- the width of each bin is 5 days
- the color of the interior of the bars is "forestgreen"
- the color outlining the bars is "darkviolet"
- the x label says "Days Closed with Ice"
- the y label says "Total"
- there is a title with words of your choosing that describe the figure

```{r problem2}
## modify this code
ggplot(monona, aes(x = duration)) +
  geom_histogram(center=70,fill="forestgreen", color="darkviolet", binwidth=5) + xlab("Days Closed with Ice") + ylab("Total") + ggtitle("Winters in Lake Monona")
```

### 3

Create a line plot with `duration` on the y axis and `year1` on the x axis. Add a blue smooth curve and a straight red line which estimate the trends.
Modify axis labels and add an appropriate title.

```{r problem3}
ggplot(monona, aes(x=year1, y=duration)) + 
  geom_line()+geom_smooth()+geom_smooth(method="lm",color="red")
```

- Compare the curve and the straight line; does the curve differ from the straight line in an important manner?

- To the nearest five days, about how long was Lake Monona closed with ice in a typical year near 1875 (i.e., what is the approximate value of the smooth curve or straight line around 1875)?

- To the nearest five days, about how long was Lake Monona closed with ice in a typical year near 2000 (i.e., what is the approximate value of the smooth curve or straight line around 2000)?



### Response

The curve does not differ in an very important way from the straight line. 

The curve does not differ in an very important way from the straight line. 

Lake Monona was closed with ice for about 90 days in a typical year near 2000. 
### 4

The next chunk of code adds columns to the `monona` data set which contain the fitted values ($\hat{y}_i$ in the notation of the lecture notes) and the residuals after fitting a straight line simple regression model to the data.
You do not need to examine or understand this code chunk.

```{r}
## Fit the model
fit = lm(duration ~ year1, data = monona)

## Add fitted values and residuals to the data set
monona = monona %>% 
  mutate(fitted = fitted(fit),
         residuals = residuals(fit))
```

Make a plot of the residuals versus year1.
Add a horizontal line with an intercept at zero. *(See example from lecture using `geom_hline()`)*.
Add a smooth curve to this residual plot as well.

```{r}
ggplot(monona, aes(x=year1, y=residuals)) + geom_point() + geom_hline(yintercept=0)+geom_smooth()
```

Using the informal interpretation discussed in lecture,
what evidence from the residual plot indicates that a straight-line simple regression model for the duration that Lake Monona is frozen versus time is an adequate model?


### Response

A straight-line simple regression model is adequate for the duration that Lake Monona is frozen versus time because it appears that there is an equal number of values above and below the line.


### 5

The next chunk of code adds a variable `winter_day` to the `monona` data set which counts the number of days after the previous June 30 on which Lake Monona first freezes.
You do not need to examine or understand this code.

```{r}
## add winter_day = #days of first freeze after June 30
monona = monona %>% 
  mutate(winter_day = as.numeric(
    first_freeze - ymd(str_c(year1,"-06-30"))))
```

Make a histogram of this variable with bins of width 5 centered
and one bin centered at 154 (which corresponds to December 1).
Select different colors to shade the interior and the boundaries of the bars.
Add appropriate axis labels and a plot title.

```{r}
## Make the histogram
ggplot(monona, aes(x=winter_day)) +
  geom_histogram(binwidth = 5, fill="cadetblue1", color="cadetblue",center=154) +xlab("Winter Days")+ylab("Frequency")+ggtitle("Lake Monona First Freeze Distribution")
```


### 6

Recreate the plot from the previous problem,
but modify scale of the x axis so that the axis is labeled with the dates Dec 1, Dec 16, Dec 31, Jan 15, Jan 30 at the values
154, 169, 184, 199, 214.
You will use the function `scale_x_continuous()` to change the scale of the x axis.
*(Hint: See an example changing the scale of the y axis in lecture 5)*.

```{r}
ggplot(monona, aes(x = winter_day)) +
  geom_histogram(binwidth = 5, center = 154, color = "darkviolet", fill = "forestgreen") +
  xlab("Days after June 30") +
  ylab("Count") +
  ggtitle("Lake Monona First Freeze Distribution",
          subtitle = "1851 - 2021") + scale_x_continuous(breaks=seq(154,214,by=15),labels = c("Dec 1", "Dec 16", "Dec 31", "Jan 15", "Jan 30"))
```

### 7

The next chunk of code reads in the exoplanet data discussed in the course notes, reduces the data set to a few variables, renames these,
and drops observations with missing masses.
You do not need to read or comment on this code chunk.

```{r}
## Note: the file begins with 100 comment lines which we skip
exo = read_csv("../../data/exoplanets_default_2021-06-15.csv", skip = 100) %>% 
  select(pl_name, hostname, discoverymethod, pl_bmasse) %>%
  rename(planet = pl_name,
         star = hostname,
         method = discoverymethod,
         mass = pl_bmasse) %>% 
  drop_na(mass)
```

The following code chunk makes a boxplot of the planet masses (units are in Earth masses, so 10 is ten times the mass of Earth) versus the method of discovery.

```{r}
ggplot(exo, aes(x = method, y = mass)) +
  geom_boxplot()
```

What are some faults that this basic plot has?

### Response

The x-labels are not readable because they are so close together. Additionally, using a boxplot is a bad idea since it isn't understandable. There is no title explaining what the graph is. 

The next code chunk adds several lines of code.
Explain the effect that each additional line highlighted with a comment does to the plot.
Try running the code with the indicated lines commented out or seek information from R Studio help, the *R for Data Science* textbook, or a Google search about each command and argument to understand what it does.

```{r}
## You may need to install the scales package if the next line causes an error
library(scales)

## Make the side-by-side boxplots
ggplot(exo, aes(x = method, y = mass)) +
  geom_boxplot() +
  ## (A) What does the following line do?
  scale_y_continuous(trans = "log10", label = comma) + ## This scales the y-axis so that it is more easily readable, and makes sure that the boxplot has easily distinguishable quartiles. 
  xlab("Method of Detection") +
  ylab("Mass (Earth masses)") +
  ## (B) What does the following line do?
  coord_flip() 
```

### Response

#### (A) effect of scale_y_continuous()
##### trans = "log10"

This transforms the axis into a log 10 scale; the values are 1, 100, 10,000, etc. 

##### label = comma

This puts commas in numbers so it's easier to read bigger values. 

#### (B) effect of coord_flip()

This exchanges the axes; the x-axis becomes the y-axis and vice-versa. 


### 8

Create a density plot of the variable `mass` from the exoplanet data set.
Use different colors for the interior of the density and the boundary line.
Describe the shape of exoplanet masses.

```{r}
ggplot(exo, aes(x=mass)) +
  geom_density(color="black", fill="purple") +
  xlab("Mass") + ylab("Frequency") + ggtitle("Mass of Exoplanets")
  
```

### Response

Most of the exoplanets are between 0 and 2500 Earth Masses. 

Make a second density plot for the base-10 logarithm of masses instead.
Describe the shape of the log-transformed exoplanet masses.

```{r}
ggplot(exo, aes(x=mass)) +
  geom_density(color="black", fill="purple") +
  xlab("Mass") + ylab("Frequency") + ggtitle("Mass of Exoplanets") + 
  scale_x_continuous(trans = "log10", label = comma)
```

### Response

There are two peaks, one at 50 earth masses and one at 2500 earth masses. 
