---
output: html_document
---

## Base R Examples

### Vectors

- Most data in the course we will create by reading data in from a text file (.txt or .csv or .tsv), from Excel files (.xlsx or .xls), or, rarely, directly from a web site using a function to read tables from HTML.
- But it is useful to know how to create a vector from scratch.
- Here are several examples.
- In each case below, the code creates a vector without assigning it to a name.
    - The vector is created, printed to output, and not saved.

#### Sequences

- Use the colon operator to create sequences of integers
- Notice in the output that each line begins with a label in square brackets which identifies the position of the first element of the vector on that line

##### The colon operator

```{r}
## 1 through 10
1:10

## 20 through 100
20:100

## 10 to 1
10:1

## Negatives
-2:5
-10:-1
```

##### seq()

- The function `seq()` allows more flexibility.
    - The elements do not need to be integers
    - The first arguments are `from` and `to`
    - The third argument is `by`, which is the step size.
    - You may also specify the desired length with `length.out`.
    - You do not need to name the arguments if they are in order
    - There are defaults (`from = 1`, `to = 1`, `by = ((to = from)/length.out))`
    
```{r}
## default sequence from 1 to 1
seq()

## sequence from 1 to 10 by 0.5
seq(1, 10, 0.5)

## same, using named arguments
seq(from = 1, to = 10, by = 0.5)

## same, if named arguments, can be passed in any order
seq(to = 10, by = 0.5, from = 1)

## same, specifying the length
seq(1, 10, length.out = 19)
```
    
##### c()

- The handy function `c()` combines arbitrary values into a vector

```{r}
c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

c(1:5, 10)

c(1, 1, 2, 3, 5, 8, 13, 21)
```

### Assignment

- The conventional symbol to assign a vector (or any other object in R) to a name is the two-character combination `<-` which looks like an arrow.

```{r}
v1 <- 1:10
v1
```

- An alternative (which I prefer, but am in the minority) is the equal sign `=` which is conventional in most computer programming languages.

```{r}
v2 = 1:10
v2
```

- Note that the equal sign has different meaning in different contexts.
    - We use this symbol to assign values to arguments in function calls, as we saw before.
    
```{r}
## The first = is assignment in creating an object
## The second = is for naming an argument
v3 = seq(1, 11, by = 2)
v3
```

- Use whichever symbol you prefer.
- It is better for readability of your code to be consistent with your choice.

### Object names

- Object names made be made up of lower case letters, `a-z`, upper case letters `A-Z`, digits `0-9`, and the symbols period `.` and under score `_`.
    - R is case sensitive, so the names `a` and `A` are different.
    - A valid name cannot begin with a number.
    - As the period has special meaning in some contexts, I tend to avoid its use.
    - A name which begins with a `.` is also good to avoid as these are conventionally reserved for special system variables and are hidden in the environment.
    - You can override the naming conventions by wrapping a name between single back quotes
    
```{r}
## trying to use an illegal name causes an error
##1a = 10

## a legal assignment
a = 10

## a legal assignment, but the variable does not appear in the environment
.a = 10

## Use the function ls() to list all objects
## Note .a is not included
ls()

## Add an argument to show the hidden variables too
## Who knew that .Random.seed was hanging around!
ls(all.names = TRUE)
```
    
### Arithmetic

- In many computer languages, you need loops to work over vectors (or arrays) of value
- In R, most of the language is *vectorized*, so you do not need explicit loops.
- When doing conventional arithmetic with vectors, all of the calculations are done element by element.

```{r}
## two small example vectors
a = 1:3
b = 4:6
a
b

## addition by element
a + b

## subtraction by element
b - a

## multiplication by element
a*b

## division by element
a/b

## exponentiation by element
a^b
```

- If you wanted to take a dot product, you could multiply by elements and then sum the results

```{r}
## dot product
sum(a*b)
```

- It is also common to do operations between a single number and a vector of numbers.
- Effectively, the single number is replicated to be the same length as the vector and then the operations occur.

```{r}
s = 5
s

## addition
a + s

## subtraction
a - s

## multiplication
a*s

## division
a/s
s/a

## exponentiation
s^a
a^s
```

#### Incompatible sizes

- R allows vector arithmetic with incompatibly-sized vectors
- The shorter vector is replicated to be the same length as the longer one.
- If it does not work evenly, then a warning is produced.

```{r}
a
d = 4:5
d

## adding a = c(1,2,3) and d = c(4,5)
## d is changed to c(4,5,4), but the last repeat of 4,5 is stopped short
## a warning is given as this was likely an error
a + d
```

## Common functions

- There are a number of common functions that act on each element of a vector individually.

```{r}
a

## exp()
exp(a)

## natural log, log()
log(a)

## log base 10
log10(a)
log(1, base = 10)

## trig functions
## Note the use of the built in value pi
## Ideal 0 may not be calculated exactly on a computer
## When a number ends e-16, this is a shorthand for times 10^(-16)
## Small numbers like this are effectively zero
sin(a * pi)
cos(a * pi)
tan(a*pi)
```

- Other functions summarize a vector of numbers into a single summary value

```{r}
a

## sum
sum(a)

## mean
mean(a)

## standard deviation
## if you have had a stats course before....
sd(a)

## length
length(a)

## maximum
max(a)

## minimum
min(a)

## median
median(a)

## quantiles
## there are many definitions of quantiles
quantile(a, 0.2)
quantile(a, 0.2, type = 3)

```

## Data types

- R has several data types
- Use the function `class()` to identify each

```{r}
## colon creates integers
a = 1:3
a
class(a)

## seq() creates numeric values
b = seq(1,3,1)
b
class(b)

## There are also character-valued vectors
x = c("A", "B", "C")
x
class(x)

## And logical
y = c(TRUE, TRUE, FALSE)
y
class(y)
```

- There are several more types
- We will see dates and times in a few weeks.

- You cannot mix data types in a vector.
- If you try, R will coerce the elements to be the same type

```{r}
d = c(1, 2, "A")

## Note that the 1 and 2 were converted to character valued
d
```

- You can change types of variables with a collection of functions with names like `as.integer()`, `as.numeric()`, `as.character()`

```{r}
b = seq(1,5)
b
class(b)

bb = as.integer(b)
bb
class(bb)

as.character(bb)
```

## Data frames

- We will mostly examine data packaged into a data frame
- Each column of a data frame is a vector
- The elements of a vector must be the same, but vectors of different types may be in the same data frame
- We will most often create data frames by reading in data from a file
- But we can make these by hand too

```{r}
## LETTERS is a built in vector with the 26 capital letters
## The square bracket operator takes a subset, here the first 10
dat = data.frame(x = LETTERS[1:10], y = 1:10)
dat
class(dat)

## Here are three ways to grab the vector that is the first column
## The $ operator
dat$x

## Double square brackets (treats the data frame as a list of its columns)
dat[[1]]
dat[['x']]

## Single brackets with a comma between dimensions
## Blank keeps all
dat[,1]
```

- We often want to know the dimensions (number of rows and columns) of a data frame

```{r}
dim(dat)
nrow(dat)
ncol(dat)
```

## Summary

- There are many base R functions for dealing with data
- We will focus moving forward on the tidyverse extension to R
- The syntax is quite different, but underneath there are a lot of base R calculations on vectors happening