\documentclass[letterpaper]{report}

\setlength\parindent{0pt}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{amsmath, amssymb}
\title{Weekly HW 006}
\author{Svadrut Kukunooru}
\date{\today}

\begin{document}
\begin{titlepage}
    \maketitle
\end{titlepage}
   \subsection{4.2.15}%
   \label{sub:4.2.15}
   Consider the differential equation $y'' - y' + 2y = 0$. A solution is a real-valued function $f$ satisfying the equation. Let $V$ be the set of all solutions to the given differential equation; define + and $\cdot$ as in Exercise 14. Prove that $V$ is a vector space.  \\ \\ 

To prove that $V$ is a vector space, we can use scalar addition and scalar multiplication to prove it. 

Say that we have two solutions to this equation, $y_1$ and $y_2$. 

\[
    (y_1 + y_2)'' - (y_1 + y_2)' + 2(y_1 + y_2) = 0
\] 
\[
    (y_1'' - y_1' + 2y_1) + (y_2'' - y_2' + 2y_2)
\] 
Since both of these mini-equations are in the same format as this equation, 
\[
0 + 0 = 0
\] 
Using scalar multiplication, 
\[
    c(y'' - y' + 2y) = 0
\] 
Since the differentiation equation already equals zero, no matter what $c$ is, the solution will always equal zero. Therefore, $V$ is a subspace and therefore a $\fbox{vector space}$.
\subsection{4.3.17}%
\label{sub:4.3.17}
Which of the following subsets of the vector space $M_{nn}$ are subspaces? \\ \\ 
(a) The set of all $n \times n$ symmetric matrices.  \\ \\ 
If $A = A^T$ and $B = B^T$ (since both are symmetric)
\[
    A^T + B^T = M_{nn}
\]
because a transposed symmetric matrix is identical to the original matrix. 
\\ \\ 
As for multiplication, multiplying a symmetric matrix by a constant will always result in the same dimensions. 
\\ \\ 
(b) The set of all $n \times  n$ diagonal matrices. \\  \\ 
All diagonal matrices are a subspace of symmetric matrices, which in turn are a subspace of the vector space $M_nn$.  \\ \\ 
(c) The set of all $n \times n$ nonsingular matrices. \\
A vector space is required to have a zero vector in it. As a zero vector can never be nonsingular, this set is not a subspace of $M_{nn}$. 

\subsection{4.4.4b}%
\label{sub:4.4.4b}
Determine whether the given vector $A$ in $M_{22}$ belongs to span $[A_1, A_2, A_3]$, where 
\[
A_1 = 
\begin{bmatrix} 
    1 & -1 \\
    0 & 3
\end{bmatrix} , 
A_2 = 
\begin{bmatrix} 
    1 & 1 \\
    0 & 2 \\ 
\end{bmatrix} , 
A_3 = 
\begin{bmatrix} 
    2 & 2 \\
    -1 & 1 
\end{bmatrix} 
\]
and 
\[
A = 
\begin{bmatrix} 
    -3 & -1 \\
    3 & 2
\end{bmatrix} 
\] 
\[
\begin{bmatrix} 
    -3 & -1 \\
    3 & 2 
\end{bmatrix} + 3A_3 = 
\begin{bmatrix} 
    3 & 5 \\
    0 & 5 
\end{bmatrix}  - 2A_2 = 
\begin{bmatrix} 
    1 & 3 \\
    0 & 1
\end{bmatrix} + A_1 = 
\begin{bmatrix} 
    2 & 2 \\
    0 & 4
\end{bmatrix} \cdot \frac{1}{2} = 
\fbox{$
\begin{bmatrix} 
    1 & 1 \\
    0 & 2 
\end{bmatrix}$}
\] 
Therefore, $A$ is in span $[A_1, A_2, A_3]$. 
\end{document}
