---
title: "Assignment 08"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

\newcommand{\E}{\mathsf{E}}
\newcommand{\Var}{\mathsf{Var}}
\renewcommand{\prob}{\mathsf{P}}

## Homework 8

#### Due Tuesday, November 16, 11:59 PM CT

Problems 1-3 are based on the random variable $X \sim \text{Binomial}(60, 0.4)$.

### Problem 1

### A

Calculate and report the mean and standard deviation of the distribution.

The mean is equal to $n \cdot p$ and the standard deviation is equal to $\sqrt{n \cdot p \cdot (1-p)}$

Therefore, the mean is 24 and the standard deviation is 3.79. 

### B

Calculate and report $\prob(X = 25)$

```{r}
dbinom(25, 60, 0.4)
```

### C

Calculate and report $\prob(X >= 25)$

```{r}
1 - pbinom(24, 60, 0.4)
```

### D

Create a graph which displays the binomial distribution with bars for $x \ge 25$ colored in a different color from the rest to help visualize your answer to **C**.

```{r}
gbinom(60, 0.4, a=0, b=60) +
  geom_binom_density(60, 0.4, a=0, b=24, color="grey")
```

### Problem 2

What is the probability that $X$ is within one, two, and three standard deviations of the mean? Round each probability to four decimal places.

```{r}
pbinom(27.79, 60, 0.4) - pbinom(20.21, 60, 0.4)
pbinom(31.58, 60, 0.4) - pbinom(16.42, 60, 0.4)
pbinom(35.37, 60, 0.4) - pbinom(12.63, 60, 0.4)
```


### Problem 3

Draw a graph of this binomial distribution with vertical lines with different line types at the end points of the intervals 1, 2, and 3 sigma from the mean.

- solid for $\mu \pm \sigma$
- dashed for $\mu \pm 2\sigma$
- dotted for $\mu \pm 3\sigma$

```{r}
gbinom(60, 0.4, a=0, b=60) + 
  geom_vline(xintercept=27.79, color = "red", linetype="solid") + 
  geom_vline(xintercept=20.21, color = "red", linetype="solid") +
  geom_vline(xintercept=31.58, color = "red", linetype="dashed") + 
  geom_vline(xintercept=16.42, color = "red", linetype="dashed") + 
  geom_vline(xintercept=35.37, color = "red", linetype="dotted") + 
  geom_vline(xintercept=12.63, color = "red", linetype="dashed")

```



### Problem 4

### A

Create a tibble where each row corresponds to a binomial distribution with the following columns:

- `n` for $n = 1000,1001, \ldots, 10,000$
- `p` equal to 0.9 for each row
- `mu` equal to the mean
- `sigma` equal to the standard deviation
- `q.975` equal to the 0.975 quantile of the distribution
- `z = (q.975 - mu)/sigma`, the standardized z-score of the quantile
- `prob` equal to $\prob(X \ge q.975)$, a right-tail probability for the distribution in each row

Display all columns and the rows of this table for which $n$ is a multiple of 1000.

```{r}
FINALQ4A <- tibble(
  n = 1000:10000, 
  p = 0.9, 
  mu = n * p, 
  sigma = sqrt(mu * (1-p)), 
  q.975 = qnorm(0.975), 
  z = (q.975 - mu)/sigma, 
  prob = qbinom(q.975, n, p)
)
FINALQ4A %>% filter(row_number() %% 1000 == 1) %>% print()
```

### B

Use a line plot to show the relationship between $n$ on the x axis and $z$ on the y axis.
Add a smooth curve to highlight the trend.
Describe the pattern you see.

```{r}
ggplot(data = FINALQ4A) + 
  geom_point(mapping = aes(x=n, y=z)) + geom_smooth(mapping = aes(x=n, y=z))
```

The z-score decreases logarithmic as the amount of trials increases. 

### C

What is the average of the values of $z$ for $n > 9000$?

```{r}
FINALQ5C <- FINALQ4A %>% filter(n>9000)
mean(FINALQ5C$z)
```

### D

Plot the right tail probability `prob` on y axis versus `n` on the x axis.

```{r}

```

### E

What is the average of the values of `prob` for $n > 9000$?

```{r}

```


### Problem 5

### A

What is the 0.975 quantile of the standard normal distribution?

```{r}

```
1.96

### B

What is the area to the right of this quantile?

0.1647

### Problem 6

For the $\text{N}(500,100)$ distribution, find values $a$ and $b$ such that $a < b$, the area between $a$ and $b$ under the normal density is equal to 0.6,
and $a$ and $b$ are equally distant from $\mu = 500$.

```{r}

```

a = 415.84, b = 584.16


### Problem 7

Graph the $\text{N}(500,100)$ distribution and shade the area between $a$ and $b$ from the previous problem.

```{r}
gnorm(mu=500, sigma=100,) + 
  geom_norm_fill(500, 100, a = 415.84, b = 584.16)
```





