---
title: "STAT 240 Discussion 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

\newcommand{\E}{\mathsf{E}}
\newcommand{\Var}{\mathsf{Var}}
\renewcommand{\prob}{\mathsf{P}}

## Questions

### 1

(a) A fair coin is tossed repeatedly until the tenth head is tossed. $X_1$ is the number of tails tossed prior to the tenth head.

REPLACE THIS TEXT WITH YOUR RESPONSE



(b) Four buckets each contain a total of five tokens each, some red and some blue. The number of red tokens in the buckets are 1, 2, 3, and 4 with blue tokens making up the remainder. One token is drawn at random from each bucket. $X_2$ is the total number of red tokens drawn.

REPLACE THIS TEXT WITH YOUR RESPONSE




(c) A random sample of students in a fitness study.
$X_3$ is the mean systolic blood pressure of the sample.

REPLACE THIS TEXT WITH YOUR RESPONSE



### 2

Let $X \sim \text{Binomial}(25, 0.4)$.

#### A

Calculate and report the mean $\mu = \E(X)$ and the standard deviation $\sigma = \sqrt{\Var(X)}$.

```{r}

```


#### B

Calculate $\prob(X \ge 14)$.

```{r}

```


### C

Create a graph of the $\text{Binomial}(25, 0.4)$ distribution.
Color the segments for $x = 14, 15, \ldots, 25$ red and use gray for the rest. Plot using `theme_minimal()`.

You may edit the code in this block which does something similar for a different problem.
Note the use of `geom_binom_density()` with different values of `a` and `b`.

```{r, fig.height = 3}
gbinom(100, 0.5, a = 30, b = 70) +
  geom_binom_density(100, 0.5, a = 35, b = 50, color = "gray") +
  geom_binom_density(100, 0.5, a = 55, b = 70, color = "red") +
  theme_minimal()
```


### 3

Create a data frame with columns `n`, `p`, `x`, `mu`, and `prob` where:

- `n` varies from 2 to 100 by twos (so each `n` value is even:  2, 4, 6, 8, ..., 100);
- `p` equals 0.5;
- `x` is `n*p`;
- `mu` is the mean of the Binomial distribution with parameters `n` and `p`;
- `prob` is $\prob(X = x)$, where $X \sim ~ \text{Binomial}(n, p)$

Show the first few rows using `head()`, then respond to the question below.


What is the relationship between `x` and `mu`?

REPLACE THIS TEXT WITH YOUR RESPONSE





### 4

The central limit theorem implies that the binomial distribution converges to a normal distribution as $n$ increases.  This problem will examine one aspect of the convergence, namely the right tail probability of being more than two standard deviations above the mean, $\prob(X > \mu + 2\sigma)$.

**(a)** What is the probability that a normal random variable with mean $\mu$ and standard deviation $\sigma$ exceeds $\mu + 2 \sigma$? Display answer rounded to four decimal places.





**(b)** 

Create a data frame with columns `n`, `p`, `mu`, `sigma`, `x`, and `prob` where:  

- `n` varies from 1 to 1000 by ones (1, 2, 3, ..., 1000);  
- `p` equals 0.5;  
- `mu` is the mean of the Binomial distribution with parameters `n` and `p`
- `sigma` is the standard deviation of the Binomial distribution with parameters `n` and `p`;
- `x = mu` + 2*sigma`;  
- `prob` is $\prob(X > x)$, where $X \sim \text{Binomial}(n, p)$

Display the first few rows of the data frame with `head()`.



**(c)** Plot the right tail probabilities versus $n$, and carryout the following modifications:

- Add a red, dashed, horizontal line at the value you found in part **(a)**.  
- Add a smooth curve which lessens the visual impact of the oscillations due to the discreteness of the binomial distribution.




### 5

Repeat problems **4B** and **4C** if $p = 0.005$.




### 5D

Ignoring the oscillations, how do the patterns of the two smooth curves in the graphs differ from one another, especially when $n$ is large?

REPLACE THIS TEXT WITH YOUR RESPONSE



### 6

Draw a graph of the binomial distribution with $(n=500, p = 0.5)$ and a graph for the binomial distribution with $(n=500, p = 0.005)$; you may use the `gbinom()` function in the *ggprob.R* script.
Scale these graphics so that the x axis is restricted to where the probabilities are relatively large; set `scale=TRUE` in the `gbinom()` function to accomplish this.  

Overlay each plot with a red normal density with a mean and standard deviation that matches the mean and standard deviation of the corresponding binomial distribution.  Consider the `geom_norm_density()` function in the *ggprob.R* script.

Compare the skewness of the distributions and comment on how this might help explain the differences in the right tail probabilities from Problem 3.

REPLACE THIS TEXT WITH YOUR RESPONSE



### 7

If $X \sim \text{N}(20, 3)$, what is $\prob(15 < X < 27)$?

```{r}

```

0.924

### 8

Draw a graph of a normal distribution with mean 20 and standard deviation 3 with the area between 15 and 27 shaded.
How does this area correspond to the previous calculated probability?

REPLACE THIS TEXT WITH YOUR RESPONSE



The shaded area equals the previous problem calculation.

### 9

Define $z_1 = (15 - 20)/3$ and $z_2 = (27 - 20)/3$.
If $Z \sim \text{N}(0,1)$, what is $\prob(z_1 < Z < z_2)$?

```{r}

```



### 10

Draw a graph of a standard normal distribution and shade the area under the curve between $z_1$ and $z_2$ from the previous problem.
How does this plotted graph compare to the graph made in question 8?

```{r}

```




