---
title: "STAT 240 Discussion 9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
source("../../scripts/viridis.R")
source("../../scripts/ggprob.R")
```

\newcommand{\E}{\mathsf{E}}
\newcommand{\Var}{\mathsf{Var}}
\renewcommand{\prob}{\mathsf{P}}

## Problems

The questions in this discussion assignment use the chimpanzee data set.
In answering these questions, you may assume the following in models for the data:

1. all trials are independent
2. trials made with a partner have the same chance of being pro-social, even if the identities of the actors or partners differ.

### Problem 1

#### A

Using the chimpanzee data,
combine all of the trials made without a partner.
Count the number of pro-social and selfish choices made.

```{r}
chimpanzee <- read_csv("../../data/chimpanzee.csv") 
FINALQ1 <- chimpanzee %>% filter(partner == "none")
sum(FINALQ1$prosocial)
sum(FINALQ1$selfish)
```
83 Pro-Social, 97 Selfish choices made.

#### B

Write a model for the number of pro-social choices.

```{r}
gbinom(180, 0.46, scale=TRUE) 
```

#### C

Use the Agresti-Coull method to construct a 95% confidence interval for the probability that a chimpanzee makes the pro-social choice in experimental conditions when there is no partner present.

```{r}

```

#### D

Interpret this confidence interval in the context of the experimental setting.

### Problem 2

Using the same data as in the previous problem,
do the following.

#### A

State hypotheses for a two-sided hypothesis test that the probability that a chimpanzee makes the pro-social choice over the long run is equal to 0.5 versus the alternative that this probability is not 0.5.


$H_0: p = 0.5 \\ 
H_a: p \neq 0.5$


#### B

Pick a test statistic, state the sampling distribution of this statistic if the null hypothesis is true,
and compute a p-value for the test.

```{r}
p_value = 1 - pbinom(83, 180, 0.5) # P(X > 49) = P(X >= 50)
p_value
```
If the null hypothesis is true, our distribution is $\text{Binomial}(180, 0.5)$. As this is almost equal to 1, we reject the null hypothesis. 

#### C

Interpret these results in the context of the experiment following the examples from lecture notes.

There is very strong evidence that the chimpanzee will make the selfish choice more than half the time in the long run under similar experimental conditions. 

### 3

Use functions from the file `ggprob.R` to visualize the calculation of the p-value by graphing an appropriate probability distribution and using differential coloring to indicate what in the plot corresponds to the p-value.
Include a title and meaningful axis labels.

```{r}

```

